### 2023-12-18 latex/Plusone_configs/paper.tex
20231215

% amsmath and amssymb packages, useful for mathematical formulas and symbols
\usepackage{amsmath,amssymb}
\usepackage{algorithm}
\usepackage{algorithmic}
\usepackage{booktabs}
\usepackage{rotating}
% Use adjustwidth environment to exceed column width (see example table in text)
\usepackage{changepage}

@@

% Text layout
\raggedright
\setlength{\parindent}{0.5cm}
\textwidth 5.25in
\textheight 8.75in

@@

\end{flushleft}
% Please keep the abstract below 300 words
\section*{Abstract}
We found that information on the VIX futures curve can guide the trading of VIX derivatives. We use two sets of features derived from the VIX futures curve to predict the next day's return rates of VIX ETFs and construct a trading strategy. In this strategy, up to 4 out of 8 different machine learning models achieved an Information Coefficient (IC) value of over 0.02 over a span of 11 years, compared to the MSE and AR(1) baseline models. This indicates that the VIX futures curve indeed has predictive power over VIX ETFs. In a strict Walk-Forward back test, a simple long-short strategy achieved an average information ratio of 1.3. We used Portfolio Overlay (PO) methods to increase the average information ratio to 1.6, demonstrating that the prediction of returns is valuable not only in relative terms but also in absolute terms.

% Please keep the Author Summary between 150 and 200 words
% Use first person. PLOS ONE authors please skip this step.
% Author Summary not valid for PLOS ONE submissions.
%\section*{Author summary}
%Lorem ipsum dolor sit amet, consectetur adipiscing elit. Curabitur eget porta erat. Morbi consectetur est vel gravida pretium. Suspendisse ut dui eu ante cursus gravida non sed sem. Nullam sapien tellus, commodo id velit id, eleifend volutpat quam. Phasellus mauris velit, dapibus finibus elementum vel, pulvinar non tellus. Nunc pellentesque pretium diam, quis maximus dolor faucibus id. Nunc convallis sodales ante, ut ullamcorper est egestas vitae. Nam sit amet enim ultrices, ultrices elit pulvinar, volutpat risus.

@@

\begin{figure}[!h]
\caption{\bf CMfs Correlations}
%\includegraphics[width=0.5\linewidth]{EDA_correlations.png}
\label{fig1}
\end{figure}

The VIX term structures describes the relationship of VIX futures contracts with varying expiration dates.The term structure is in contango when VIX futures with longer maturities have higher prices than futures with shorter maturities, backwardation occurs when prices of VIX futures with shorter time to expiration exceed those with longer maturities. Our empirical analysis,  as showcased in exhibit 1 through K-means clustering, robustly indicates a dominant contango shape within the VIX future term structure. Such phenomenon can be partially explained based on volatility risk premium theory (Carr \& Wu, 2008) and  Johnson (2017) work, which divides VIX term strucuture  into two fundamental components: conditional volatility expectations and risk premiums. It becomes apparent that investors stipulate a larger risk premium for further VIX options, as contrasted with the shorter-maturity VIX options except during certain financial crisis. Notably, the VIX futures term structure holds predictive power when anomalies in the term structure revert to the classical shape given strong mean-reverting property of VIX futures illustrated by Avellaneda and Papanicolaou (2019).
The literature has extensively explored predictive power of implied volatility (IV),  IV term structure and VIX term structure using classical statistical models. Wang and Wang (2016) among others find that incorporating IV can enhance forecast accuracy within the GARCH model framework. Additionally, studies by Bush et al. (2011), Byun and Kim (2013), and Haugom et al. (2014) have demonstrated the informative power of IV using the heterogeneous autoregressive (HAR) model.transitioning towards term structures, few studies confirms the prediction power of term structures for various financial instruments. Chang (2016) shows the information gained in VIX term structure significantly aid in predicting S\&P 500 index returns volatility while Clements (2020) uses HAR model to further affirm that the information derived from IV term structure possesses exceptional predictive power for forecasting both the level and directional changes in stock return volatility, even during the volatile period of the Global Financial Crisis.Chen (2022) employes CHH model and confirms that the volatility factor constructed from swaptions IV term structures holds significant predictive power on excess bond returns. Lastly, Ornelas and Mauad (2019) provides empirical evidence of the IV term structures informative power in forecasting exchange rate returns. Based on the robust findings from above research use classical statistical models to exploit the predictive power of term structures, in this paper, we use HAR model as the baseline to compare with other selected machine approaches.

Despite the booming of machine learning (ML) approaches, relatively scarce studies use machine learning approach to exploit VIX-related issues. Hosker et al (2018) finds that recurrent neural network(RNN) and long short-term memory(LSTM) shows the best performance for forecasting 1-month VIX futures 3-days and 5-days ahead among principal components analysis and ARIMA model. Hirsa and et al. (2021) apply neural network models (random forest, support vector machines, feed-forward neural network and LSTM) to investigate the approaches that could replicate VIX index and VIX futures with fewer options compared to CBOE original methodology. Vrontos (2021) applies various machine learning models and discovers that ML models are statistical and economic superior compared to standard econometric models when forecasting the directional changes of VIX index. Lastly, Avellaneda (2021) generates VIX futures trading signals by including functional dependence of VIX term structures, VIX futures position as well as expected utility in neural network and finds that this approach shows better portfolio performance in the out of sample backtesting.

In this paper, we use multiple types of machine learning models which includes Recurrent neural network models, tree-based model, transformer and graph attention neural network to exploit informative power of VIX futures term structure on next-day return of  VIX constant maturity instruments.

By working closely with Avellaneda (2021) work,we first use information extracted from the VIX term structure to construct various VIX term-structure-related features and combine major financial market features, we then implement various machine learning methods to predict the next day return of six VIX constant-maturity instruments. Lastly, to evaluate the economic performances of our work, we utilise machine learning prediction results as trading signals to implement two trading strategies: The long-short strategy and mean-variance-adjusted portfolio optimization strategy. Our results reveal that the VIX term structure has superior informative power in predicting the return of VIX constant-maturity instruments.

The key contributions of this study can be summarised as follows, first, based on the VIX futures term structures decomposition approach proposed by Avellaneda(2021), features with term-structure information, personal information, and major market trends information are included in the datasets. Second, various state-of-art machine learning approaches are introduced to predict the VIX rolling-ETFs next-day returns. Third, two trading strategies are used in the backtesting which are, the long-short strategy based on machine learning ranking and a mean-variance adjusted portfolio optimization strategy.
The results reveal that firstly, the VIX futures curve has the ability to predict the next-day return of VIX ETFs. Thirdly, The economic performance of machine learning models relying on the adjusted mean-variance portfolio optimization strategy generally outperforms that of models based on the long-short strategy. The remainder of the paper is composed as follows: Section 2 provides proposed research methodology(1.VIX term structures decompositions 2.machine learning models) section 3 describes data and experiment design, Section 4 analysis results, Section 5 concludes and discuss, Appendix: graph and code.

\section*{Research Methodologies and Data}
\subsection*{1. Trading signals constructions}
The VIX is calculated by the Chicago Board Options Exchange (CBOE) and is based on the prices of options on the S\&P 500 Index.
\\Let $t$ denote time and let $VIX_t$ denote the value of VIX on that date. At time $t$,
we can view lots of VIX futures contracts with different expiration dates:
\begin{eqnarray}
    {F^i_t} := \text{VIX futures expiring at time $T_i$ at time t}
\end{eqnarray}
where $i$ denote the index of the VIX futures contracts with a ascending expiration dates $T_1 < T_2 < ... < T_d$.
\\A term-structure of constant-maturity VIX futures(CMFs), consist of 1,2,...,6 months in our case, are constructed as
a linear interpolation of two nearest expiration dates VIX futures:
\begin{eqnarray}
\label{eq:defineV}
    {V^j_t} := \omega^j_tF^{L_j}_t\ + (1 - \omega^j_t)F^{R_j}_t \text{for j in \{1,2,...,6\}}
\end{eqnarray}
where $L_j$ represents the VIX futures of the nearest expiration dates close to $j$-month on the left of time axis,
$R_j$ represents the VIX futures of the nearest expiration dates close to $j$-month on the right of time axis, and
$\omega^j_t = \frac{T_{R_j} - 30j}{T_{R_j} - T_{L_j}}$. Note that $VIX_t$ is like a zero-horizon CMF.
CMFs are commonly used for they do not suffer fluctuations caused by contract expiry.

\subsubsection*{Rolling VIX Futures Derivatives and Strategies}
Through a rolling VIX futures derivatives or strategies can reappearance the daily returns of CMFs.
SPVXSP, SPVIX2ME, SPVIX3ME, SPVIX4ME, SPVXMP, SPVIX6ME represents 1-6 months CMFs, they can be traded or replicated with trades in the rolling VIX futures strategies.
Derivatives like ETFs and ETNs or Strategies maintains the CMF weights of equation \ref{eq:define V} for fixed maturity $j-month$. For each $j$, we denote $R^j$
the daily return of rolling VIX futures strategy, trading cost is not considered:
\begin{eqnarray}
    {R^j_t} = \frac{\omega^j_t{\Delta}F^{L_j}_t\ + (1 - \omega^j_t){\Delta}F^{R_j}_t}{\omega^j_tF^{L_j}_t\ + (1 - \omega^j_t)F^{R_j}_t}
\end{eqnarray}
where${\Delta}F^{L_j}_t = F^{L_j}_{t+1} - F^{L_j}_t$. We denote ${\Delta}t = \frac{1}{252}$, transform equation in terms of the CMFs:
\begin{eqnarray}
\label{eq: $R^j_t$2}
    {R^j_t} = \frac{{\Delta}V^j_t}{V^j_t} - \frac{F^{R_j}_{t+1} - F^{L_j}_{t+1}}{V^j_t(T_{R_j} - T_{L_j}){\Delta}t}{\Delta}t
\end{eqnarray}
Let Roll denote the drift term in equation \ref{eq: $R^j_t$2}, it referred to as the roll yield of rolling VIX futures strategy:
\begin{eqnarray}
    {Roll^j_{t+1}} := - \frac{F^{R_j}_{t+1} - F^{L_j}_{t+1}}{V^j_t(T_{R_j} - T_{L_j}){\Delta}t}
\end{eqnarray}
In a Contango circumstances, $F^{R_j}_{t+1} > F^{L_j}_{t+1}$ means $Roll < 0$. On the opposite, in a Backwardation circumstances $Roll > 0$.
We re-write equation as follows:
\begin{eqnarray}
\label{eq: $R^j_t$3}
    {R^j_t} = \frac{{\Delta}V^j_t}{V^j_t} + Roll^j_{t+1}{\Delta}t
\end{eqnarray}
From equation \ref{eq: $R^j_t$3} we see $R^j_t$ consist of two parts, the change in $V^j_t$ and roll yield. As history data shown, the most likely VIX futures
curves are contango, so the value of the rolling VIX futures strategies decay.
\\xxx shows that volatility is a mean-reversion process, in a ideal world, suppose $V^j_t$ and CMFs are both stable for a long time like a month, then
the return of $V^j_t$ should be $V^j_t - V^{j-1}_t$, so we denote $\mu$ as :
\begin{eqnarray}
\label{eq: mu}
    {\mu^j_t} = V^j_t - V^{j-1}_t
\end{eqnarray}
$\mu^j_t$ contains instantaneous market views of return about $V^j_t$.
ynorm

@@

\subsubsection*{2.1 XGBoost}
 The concept of boosting, proposed by Freund and Schapire (1999), revolves around an ensemble model strategy that combines multiple weak learners to create a more robust learner. Friedman (1999) introduced Gradient Tree Boosting, an improvement in boosting that incorporates the gradient descent algorithm into the loss function to minimize errors. Expanding on gradient boosting, Chen and Guestrin (2016) introduce XGBoost, a scalable end-to-end tree boosting system based on additive training strategy that augments gradient tree boosting with regularization, efficient handling of sparse data, and enhanced computational efficiency, resulting in superior speed and accuracy. Eq.1 denotes the objective function of XGBoost where \(\mathbf{\sum_{i=1}^{n}l(y_i,\hat{y}_i)}\) represents the training loss function and \(\mathbf{{\sum_{k=1}^{K}}\Omega(f_k)}\) denotes the complexity of trees.
\begin{eqnarray}
\label{eq:schemeP}
\mathrm{Obj}=\underbrace{\mathbf{\sum_{i=1}^{n}l(y_i,\hat{y}_i)}}_{Training \: Loss} +\underbrace{\mathbf{{\sum_{k=1}^{K}}\Omega(f_k)}}_{Complexity \: of \: trees}
\end{eqnarray}

\subsubsection*{2.2 Multilayer Perceptron (MLP)}

(introduction of MLP).
In our experiment, we use leaky ReLU introduced by Mass (2013) as the activation function \(\phi\) for MLP, the mathematical expression of leaky ReLU  is shown in Eq.2. Given activation function \(\phi\), features \(X \in R^ {n\times d}\) where \(x\) presents the batch size and \(d\) presents the dimensionalities
\begin{eqnarray}
\label{eq:schemeP}
\mathrm{\phi^{(i)}} =\mathbf{max({W^{(i)}}^{T}x,0)}=
\begin{cases}
\mathbf{{W^{(i)}}^{T}x},\quad &\mathbf{{W^{(i)}}^{T}x}>0\\0, \quad else
\end{cases}
\end{eqnarray}
where \(W^{(i)}\) is the weight vector of \(i_{th}\) hidden unit while \(x\) is the input features. A naive MLP can be illustrated as fig.\ref{MLP}
\begin{figure}[!h]
%\includegraphics[width=0.5\linewidth]{MLP.jpg}
\caption{{\bf Naive MLP}}
\label{MLP}
\end{figure}

\subsubsection*{2.3 Recurrent Neural Network (RNN)}
We then implement multiple machine learning models based on RNN architecture which aims to find patterns in sequences of data. The work of Schmidt (2019) provides a comprehensive overview of the basic architecture and functions of RNNs while \textit{Sherstinsky (2023) work focuses on deriving the classical RNN formulation from differential equations and addresses the challenges in training standard RNNs. The transformation of RNN into the "Vanilla LSTM" network is explained through logical arguments, with a detailed description of the LSTM system's equations and entities. }
RNN train a function $f$:
\begin{eqnarray}
\label{eq: RNN}
    {h_{t}, y_t} = f(h_{t - 1}, x_t)
\end{eqnarray}
where $h$ represents hidden layer, it carry sequence information by $t - 1$, $x_t$ is the input at $t$, $f$ outputs result $y_t$ at $t$ and sequence information $h_t$.

@@

Eq \ref {eq: LSTM} and Eq \ref {eq: ALSTM} illustrates the mathematical details for the implementation of LSTM and ALSTM in our study.
LSTM framework denote $c$ as the long-term sequence information distinguish from $h$, and design gates to filter information. A gate cell can be represent as:
\begin{eqnarray}
\label{eq: Gate}
    g = \sigma(W\bullet[h_{t - 1}, x_t] + b)
\end{eqnarray}
where $W, b$ are trainable parameters, $\sigma$ represent activation function, so $g$ can be use as a gate. LSTM framework design three gate on RNN as follow:
\begin{eqnarray}
\label{eq: LSTM}
    \begin {cases} c_t = g^f{\odot}c_{t - 1} - g^i{\odot}tanh(W\bullet[h_{t - 1}, x_t] + b)
    \\h^t = g^o{\odot}tanh(c_t)
    \\y^t = f(h_t)
    \end {cases}
\end{eqnarray}
$g^f, g^i, g^o$ represent three gates, and $f$ is some function link LSTM outputs to downstream task.
\\ALSTM add a encoder and a decoder attention-based layers in LSTM, Given $n$ series, stage one encoder layer can learn attention at a same time and transform to new input.
\begin{eqnarray}
\label{eq: ALSTM1}
    \begin {cases} e^k_t = tanh(W\bullet[h_{t - 1}, c_{t - 1}] + bx^k)
    \\{\alpha}^k_t = \frac{exp(e^k_t)}{\sum_{i=1}^n exp(e^i_t)}
    \\\widetilde{x_t} = ({\alpha^1_t}x^1_t, {\alpha^2_t}x^2_t,..., {\alpha^n_t}x^n_t)
    \end {cases}
\end{eqnarray}
with $\widetilde{x_t}$ as input, stage one construct with a LSTM framework named $L_1$, stage two decoder layer learn attention with all the hidden state $h_t$ of $L_1$,
the output $\widetilde{h_t}$ transform the label $y$ of another LSTM framework named $L_2$, so two attention layers can learn relations between $n$ series.
Let $d$ denote the hidden state of $L_2$, transform of $y$ can be formulate as:
\begin{eqnarray}
\label{eq: ALSTM2}
    \begin {cases} l_t = tanh(W\bullet[d_{t - 1}, c^2_{t - 1}] + bh^i)
    \\{\beta}_t = \frac{exp(l_t)}{\sum_{i=1}^T exp({\beta}_t)}
    \\c_t = \sum_{i=1}^T {\beta}_ih_i
    \\\widetilde{y_{t - 1}} = W\bullet[y_{t - 1}; c_{t - 1}] + b
    \end {cases}
\end{eqnarray}
$\tilde{y}$ is the input of $L_2$.

@@

\subsubsection*{2.4 Temporal Covolutional Networks}
Temporal covolutional networks (TCN), originally introduced by Lea et al. (2016), is a type of neural network that employs convolutional layers to capture dependencies in sequential data. Unlike traditional RNN and LSTMs, TCNs use 1D convolutions to process input sequences in parallel, making them highly parallelizable and capable of capturing long-range dependencies. Implementation details can be found in eq.5.

\begin{eqnarray}
\label{eq: TCN}
    TCN = FCN + \text{causal convolutions}
\end{eqnarray}
where FCN is fully-convolutional network with one dimension.

>>>>

% amsmath and amssymb packages, useful for mathematical formulas and symbols
\usepackage{amsmath,amssymb}
\usepackage{algorithm}
\usepackage{algorithmic}

% Use adjustwidth environment to exceed column width (see example table in text)
\usepackage{changepage}

@@

% Text layout
\raggedright
\setlength{\parindent}{0.5cm}
\textwidth 5.25in
\textheight 9 in

@@

\end{flushleft}
% Please keep the abstract below 300 words
\section*{Abstract}
This study investigates the predictive capabilities of term structures derived from VIX constant-maturity futures (VIX CMFs), expanding from the methodology proposed by Avellaneda (2021). Both predictive and back testing performances are evaluated based on various metrics with a strict walk-forward expanding time windows procedure. The main conclusions are the VIX Cmfs term structure information, specifically \(\mu_t\) and \(\triangle roll\), extracted from \(V_t\) and \(roll_t\), is highly effective in predicting the next-day returns of VIX CMFs. This finding is supported by back-testing analysis, indicating potential economic benefits. However, it was also observed that time-series derivatives have a minimal impact on enhancing predictive accuracy. \textit{(machine learning)}. Another contribution of our work is the development of the Constrained-Mean-Variance Portfolio Optimization (C-MVO) strategy. This strategy demonstrates superior back-testing performance across most models compared to the ordinal-based long-short strategy, offering valuable insights and practical implications for the development of trading strategies for VIX CMFs based on numerical predictions.
% Please keep the Author Summary between 150 and 200 words
% Use first person. PLOS ONE authors please skip this step.
% Author Summary not valid for PLOS ONE submissions.
%\section*{Author summary}
%Lorem ipsum dolor sit amet, consectetur adipiscing elit. Curabitur eget porta erat. Morbi consectetur est vel gravida pretium. Suspendisse ut dui eu ante cursus gravida non sed sem. Nullam sapien tellus, commodo id velit id, eleifend volutpat quam. Phasellus mauris velit, dapibus finibus elementum vel, pulvinar non tellus. Nunc pellentesque pretium diam, quis maximus dolor faucibus id. Nunc convallis sodales ante, ut ullamcorper est egestas vitae. Nam sit amet enim ultrices, ultrices elit pulvinar, volutpat risus.

@@

\begin{figure}[!h]
\caption{\bf CMFs Correlations}
%\includegraphics[width=0.5\linewidth]{EDA_correlations.png}
\label{fig1}
\end{figure}

The VIX term structures describes the relationship of VIX futures contracts with varying expiration dates.The term structure is in contango when VIX futures with longer maturities have higher prices than futures with shorter maturities, backwardation occurs when prices of VIX futures with shorter time to expiration exceed those with longer maturities. Our empirical analysis,  as showcased in exhibit 1 through K-means clustering, robustly indicates a dominant contango shape within the VIX future term structure. Such phenomenon can be partially explained based on volatility risk premium theory (Carr \& Wu, 2008) and  Johnson (2017) work, which divides VIX term strucuture  into two fundamental components: conditional volatility expectations and risk premiums. It becomes apparent that investors stipulate a larger risk premium for further VIX options, as contrasted with the shorter-maturity VIX options except during certain financial crisis. Notably, the VIX futures term structure holds predictive power when anomalies in the term structure revert to the classical shape given strong mean-reverting property of VIX futures illustrated by Avellaneda and Papanicolaou (2019).
\\The literature has extensively explored predictive power of implied volatility (IV),  IV term structure and VIX term structure using classical statistical models. Wang and Wang (2016) among others find that incorporating IV can enhance forecast accuracy within the GARCH model framework. Additionally, studies by Bush et al. (2011), Byun and Kim (2013), and Haugom et al. (2014) have demonstrated the informative power of IV using the heterogeneous autoregressive (HAR) model.transitioning towards term structures, few studies confirms the prediction power of term structures for various financial instruments. Chang (2016) shows the information gained in VIX term structure significantly aid in predicting S\&P 500 index returns volatility while Clements (2020) uses HAR model to further affirm that the information derived from IV term structure possesses exceptional predictive power for forecasting both the level and directional changes in stock return volatility, even during the volatile period of the Global Financial Crisis.Chen (2022) employes CHH model and confirms that the volatility factor constructed from swaptions IV term structures holds significant predictive power on excess bond returns. Lastly, Ornelas and Mauad (2019) provides empirical evidence of the IV term structures informative power in forecasting exchange rate returns. Based on the robust findings from above research use classical statistical models to exploit the predictive power of term structures, in this paper, we use HAR model as the baseline to compare with other selected machine approaches.
\\Despite the booming of machine learning (ML) approaches, relatively scarce studies use machine learning approach to exploit VIX-related issues. Hosker et al (2018) finds that recurrent neural network(RNN) and long short-term memory(LSTM) shows the best performance for forecasting 1-month VIX futures 3-days and 5-days ahead among principal components analysis and ARIMA model. Hirsa and et al. (2021) apply neural network models (random forest, support vector machines, feed-forward neural network and LSTM) to investigate the approaches that could replicate VIX index and VIX futures with fewer options compared to CBOE original methodology. Vrontos (2021) applies various machine learning models and discovers that ML models are statistical and economic superior compared to standard econometric models when forecasting the directional changes of VIX index. Lastly, Avellaneda (2021) generates VIX futures trading signals by including functional dependence of VIX term structures, VIX futures position as well as expected utility in neural network and finds that this approach shows better portfolio performance in the out of sample backtesting.
\\This paper employs a variety of machine learning models, including neural-based network models, tree-based models, and vanilla linear regression model to explore the informative power of VIX CMFs term structure on next-day returns of VIX constant maturity instruments. By working closely with Avellaneda (2021) work,we first use variables extracted from the VIX term structure to construct various VIX term-structure-related features and combine major financial market features, we then implement various machine learning methods to predict the next day return of six VIX CMFs. To asses the economic impact among models, we utilise machine learning prediction results as trading signals, implementing two trading strategies: The long-short strategy and constrained mean-variance optimization(C-MVO) strategy.
\\The key contributions of this study can be summarised as follows,  firstly, leveraging the VIX CMFs decomposition methods proposed by Avellaneda (2021), we demonstrate that the term structure information \(\mu_t\) and \(\triangle roll\) extracted from \(V_t\) and \(roll_t\) is highly informative for predicting next-day returns of VIX CMFs, and could bring economic benefits based on our back-testing analysis. However, we also found that time-series derivatives contribute minimally to predictive enhancements. \textit{Second, various state-of-art machine learning approaches are introduced to predict the VIX rolling-ETFs next-day returns.} Thirdly, our proposed Constrained-Mean-Variance Portfolio Optimization (C-MVO) strategy exhibits superior back-testing performance across most models compared to the long-short strategy, providing valuable insights for researchers and practitioners looking to develop VIX CMFs trading strategies by following the path of numerical predictions. The remainder of the paper is organised as follows: Section 2 details the decomposition methods of CMFs term structure and the machine learning approaches employed.  Section 3 describes data preparation and experiment designs, Section 4 analyse and discuss results, Appendix includes supporting graphs and tables.

\section*{Research Methodologies and Data}
\subsection*{1. Trading signals constructions}
The VIX is calculated by the Chicago Board Options Exchange (CBOE) and is based on the prices of options on the S\&P 500 Index.
\\Let $t$ denote time and let $VIX_t$ denote the value of VIX on that date. At time $t$,
we can view lots of VIX futures contracts with different expiration dates:
\begin{eqnarray}
    \mathrm{F^i_t} := \text{VIX futures expiring at time $T_i$ at time t}
\end{eqnarray}
where $i$ denote the index of the VIX futures contracts with a ascending expiration dates $T_1 < T_2 < ... < T_d$.
\\A term-structure of constant-maturity VIX futures(CMFs), consist of \({1,2,...,6}\) months in our case, are constructed as
a linear interpolation of two nearest expiration dates VIX futures:
\begin{eqnarray}
\label{eq:defineV}
    \mathrm{V^j_t} := \omega^j_tF^{L_j}_t\ + (1 - \omega^j_t)F^{R_j}_t \text{for j in \{1,2,...,6\}}
\end{eqnarray}
where $L_j$ represents the VIX futures of the nearest expiration dates close to $j$-month but with sooner expiration period than $j$-month,
$R_j$ represents the VIX futures of the nearest expiration dates close to $j$-month but with later expiration period than $j$-month, and
$\omega^j_t = \frac{T_{R_j} - 30j}{T_{R_j} - T_{L_j}}$. Note that $VIX_t$ is like a zero-horizon CMF.
CMFs are commonly used for they do not suffer fluctuations caused by contract expiry.

\subsubsection*{Rolling VIX Futures Derivatives and Strategies}
By using rolling VIX futures derivatives or strategies, we can replicate the daily returns of CMFs.
SPVXSP, SPVIX2ME, SPVIX3ME, SPVIX4ME, SPVXMP, SPVIX6ME represents 1-6 months CMFs, they can either be traded with published financial derivatives or replicated with rolling VIX futures strategies.
Derivatives like ETFs and ETNs or Strategies maintains the CMF weights of equation ~\ref{eq:defineV} for fixed maturity $j-month$. For each $j$, we denote $R^j$
as the daily return of rolling VIX futures strategy, trading cost is not considered:
\begin{eqnarray}
    \mathrm{R^j_t} = \frac{\omega^j_t{\Delta}F^{L_j}_t\ + (1 - \omega^j_t){\Delta}F^{R_j}_t}{\omega^j_tF^{L_j}_t\ + (1 - \omega^j_t)F^{R_j}_t}
\end{eqnarray}
where${\Delta}F^{L_j}_t = F^{L_j}_{t+1} - F^{L_j}_t$. We denote ${\Delta}t = \frac{1}{252}$, transform equation in terms of the CMFs:
\begin{eqnarray}
\label{eq: $R^j_t$2}
    \mathrm{R^j_t} = \frac{{\Delta}V^j_t}{V^j_t} - \frac{F^{R_j}_{t+1} - F^{L_j}_{t+1}}{V^j_t(T_{R_j} - T_{L_j}){\Delta}t}{\Delta}t
\end{eqnarray}
Let Roll denote the drift term in equation \ref{eq: $R^j_t$2}, it referred to as the roll yield of rolling VIX futures strategy:
\begin{eqnarray}
\label{roll}
    \mathrm{Roll^j_{t+1}} := - \frac{F^{R_j}_{t+1} - F^{L_j}_{t+1}}{V^j_t(T_{R_j} - T_{L_j}){\Delta}t}
\end{eqnarray}
In a Contango circumstances, $F^{R_j}_{t+1} > F^{L_j}_{t+1}$ means $Roll < 0$. On the opposite, in a Backwardation circumstances $Roll > 0$.
We re-write equation as follows:
\begin{eqnarray}
\label{eq: $R^j_t$3}
    \mathrm{R^j_t} = \frac{{\Delta}V^j_t}{V^j_t} + Roll^j_{t}{\Delta}t
\end{eqnarray}
From equation \ref{eq: $R^j_t$3} we see $R^j_t$ consist of two parts, the change in $V^j_t$ and roll yield. As history data shown, the most likely VIX futures
curves are contango, so the value of the rolling VIX futures strategies decay.
\\Avellaneda(2019) shows that volatility is a mean-reversion process, in an ideal world, suppose $V^j_t$ and CMFs are both stable for long-period such as one month, then
the return of $V^j_t$ should be $V^j_t - V^{j-1}_t$, so we denote $\mu$ as :
\begin{eqnarray}
\label{eq: mu}
    \mathrm{\mu^j_t} = V^j_t - V^{j-1}_t
\end{eqnarray}
$\mu^j_t$ contains instantaneous market views of return about $V^j_t$.

@@

\subsubsection*{2.1 XGBoost}
 The concept of boosting, proposed by Freund and Schapire (1999), revolves around an ensemble model strategy that combines multiple weak learners to create a more robust learner. Friedman (1999) introduced Gradient Tree Boosting, an improvement in boosting that incorporates the gradient descent algorithm into the loss function to minimize errors. Expanding on gradient boosting, Chen and Guestrin (2016) introduce XGBoost, a scalable end-to-end tree boosting system based on additive training strategy that augments gradient tree boosting with regularization, efficient handling of sparse data, and enhanced computational efficiency, resulting in superior speed and accuracy. Eq.\ref{eq:schemeP} denotes the objective function of XGBoost where \(\mathbf{\sum_{i=1}^{n}l(y_i,\hat{y}_i)}\) represents the training loss function and \(\mathbf{{\sum_{k=1}^{K}}\Omega(f_k)}\) denotes the complexity of trees.
\begin{eqnarray}
\label{eq:schemeP}
\mathrm{Obj}=\underbrace{\mathbf{\sum_{i=1}^{n}l(y_i,\hat{y}_i)}}_{Training \: Loss} +\underbrace{\mathbf{{\sum_{k=1}^{K}}\Omega(f_k)}}_{Complexity \: of \: trees}
\end{eqnarray}

\subsubsection*{2.2 Multilayer Perceptron (MLP)}
The Multi-Layer Perceptron (MLP) is a class of feed-forward artificial neural network (ANN). It consists of at least three layers of nodes: an input layer, a hidden layer, and an output layer. Unlike a linear perceptron, each node (or neuron) in the MLP uses a nonlinear activation function. This aspect of MLP allows it to distinguish data that is not linearly separable, which is a limitation of the basic perceptron. MLPs are widely used for solving complex problems in machine learning, including pattern recognition and classification tasks, by learning a non-linear function approximator for either classification or regression. They are trained using back propagation, ensuring that the error between the actual and predicted outputs is minimized. Due to their ability to capture complex relationships in data, MLPs have become a fundamental tool in the field of deep learning and neural networks.
In our experiment, we use leaky ReLU introduced by Mass (2013) as the activation function \(\phi\) for MLP, the mathematical expression of leaky ReLU  is shown in Eq.2. Given activation function \(\phi\), features \(X \in R^ {n\times d}\) where \(x\) presents the batch size and \(d\) presents the dimensionalities
\begin{eqnarray}
\label{eq:schemeP}
\mathrm{\phi^{(i)}} =\mathbf{max({W^{(i)}}^{T}x,0)}=
\begin{cases}
\mathbf{{W^{(i)}}^{T}x},\quad &\mathbf{{W^{(i)}}^{T}x}>0\\0, \quad else
\end{cases}
\end{eqnarray}
where \(W^{(i)}\) is the weight vector of \(i_{th}\) hidden unit while \(x\) is the input features. A naive MLP can be illustrated as fig.\ref{MLP}
\begin{figure}[!h]
%\includegraphics[width=0.5\linewidth]{MLP.jpg}
\caption{{\bf Naive MLP}}
\label{MLP}
\end{figure}

\subsubsection*{2.3 Recurrent Neural Network (RNN)}
We then implement multiple machine learning models based on RNN architecture which aims to find patterns in sequences of data. The work of Schmidt (2019) provides a comprehensive overview of the basic architecture and functions of RNNs while Sherstinsky (2023) work focuses on deriving the classical RNN formulation from differential equations and addresses the challenges in training standard RNNs. The transformation of RNN into the "Vanilla LSTM" network is explained through logical arguments, with a detailed description of the LSTM system's equations and entities.
RNN train a function $f$:
\begin{eqnarray}
\label{eq: RNN}
    {h_{t}, y_t} = f(h_{t - 1}, x_t)
\end{eqnarray}
where $h$ represents hidden layer, it carry sequence information by $t - 1$, $x_t$ is the input at $t$, $f$ outputs result $y_t$ at $t$ and sequence information $h_t$.

@@

Eq \ref {eq: LSTM} and Eq \ref {eq: ALSTM1},\ref {eq: ALSTM2} illustrates the mathematical details for the implementation of LSTM and ALSTM in our study.
LSTM framework denote $c$ as the long-term sequence information distinguish from $h$, and design gates to filter information. A gate cell can be represent as:
\begin{eqnarray}
\label{eq: Gate}
    g = \sigma(W\bullet[h_{t - 1}, x_t] + b)
\end{eqnarray}
where $W, b$ are trainable parameters, $\sigma$ represent activation function, so $g$ can be use as a gate. LSTM framework design three gate on RNN as follow:
\begin{eqnarray}
\label{eq: LSTM}
    \begin {cases} c_t = g^f{\odot}c_{t - 1} - g^i{\odot}tanh(W\bullet[h_{t - 1}, x_t] + b)
    \\h^t = g^o{\odot}tanh(c_t)
    \\y^t = f(h_t)
    \end {cases}
\end{eqnarray}
$g^f, g^i, g^o$ represent three gates, and $f$ is some function link LSTM outputs to downstream task.
\\ALSTM add a encoder and a decoder attention-based layers in LSTM, Given $n$ series, stage one encoder layer can learn attention at a same time and transform to new input.
\begin{eqnarray}
\label{eq: ALSTM1}
    \begin {cases} e^k_t = tanh(W\bullet[h_{t - 1}, c_{t - 1}] + bx^k)
    \\{\alpha}^k_t = \frac{exp(e^k_t)}{\sum_{i=1}^n exp(e^i_t)}
    \\\widetilde{x_t} = ({\alpha^1_t}x^1_t, {\alpha^2_t}x^2_t,..., {\alpha^n_t}x^n_t)
    \end {cases}
\end{eqnarray}
with $\widetilde{x_t}$ as input, stage one construct with a LSTM framework named $L_1$, stage two decoder layer learn attention with all the hidden state $h_t$ of $L_1$,
the output $\widetilde{h_t}$ transform the label $y$ of another LSTM framework named $L_2$, so two attention layers can learn relations between $n$ series.
Let $d$ denote the hidden state of $L_2$, transform of $y$ can be formulate as:
\begin{eqnarray}
\label{eq: ALSTM2}
    \begin {cases} l_t = tanh(W\bullet[d_{t - 1}, c^2_{t - 1}] + bh^i)
    \\{\beta}_t = \frac{exp(l_t)}{\sum_{i=1}^T exp({\beta}_t)}
    \\c_t = \sum_{i=1}^T {\beta}_ih_i
    \\\widetilde{y_{t - 1}} = W\bullet[y_{t - 1}; c_{t - 1}] + b
    \end {cases}
\end{eqnarray}
$\tilde{y}$ is the input of $L_2$.

@@

%\subsubsection*{2.4 Temporal Covolutional Networks}
% Temporal covolutional networks (TCN), originally introduced by Lea et al. (2016), is a type of neural network that employs convolutional layers to capture dependencies in sequential data. Unlike traditional RNN and LSTMs, TCNs use 1D convolutions to process input sequences in parallel, making them highly parallelizable and capable of capturing long-range dependencies. Implementation details can be found in eq.5.

% \begin{eqnarray}
% \label{eq: TCN}
%     TCN = FCN + \text{causal convolutions}
% \end{eqnarray}
% where FCN is fully-convolutional network with one dimension.
